---
---
---

# R for Data Science: Complete Study Plan & Syllabus

### Collaborative Book Review Learning Group \| 24 Chapters \| 25 Weeks

**Developed by**: Gideon J I, Junior Research Fellow\
**Institution**: Christian Medical College, Vellore\
**Email**: [gideon.ji.inst\@cmcvellore.ac.in](mailto:gideon.ji.inst@cmcvellore.ac.in){.email}\
**Date Created**: January 29, 2026

------------------------------------------------------------------------

## 📋 OVERVIEW

**Objective**: Master R for data science through collaborative weekly book reviews\
**Primary Resource**: *R for Data Science (1st Edition)* by Hadley Wickham & Garrett Grolemund\
**Duration**: 25 weeks (February 4 – July 22, 2026)\
**Format**: 60-minute weekly meetups with rotating leadership\
**Group Size**: 14 participants\
**Level**: Beginner to Intermediate

------------------------------------------------------------------------

## 🎯 LEARNING GOALS

By the end of this course, participants will be able to:

1.  **Import & wrangle data** – Load, clean, reshape messy datasets into analysis-ready format
2.  **Transform & explore** – Use dplyr to filter, summarize, and join datasets; conduct EDA
3.  **Visualize effectively** – Create exploratory and publication-ready plots with ggplot2
4.  **Write functions & automate** – Build reusable functions; iterate efficiently with purrr
5.  **Model data** – Fit simple and complex models to understand relationships
6.  **Communicate results** – Generate reproducible reports with R Markdown
7.  **Apply to epidemiology** – Use all skills on real STH surveillance and research data

------------------------------------------------------------------------

## 📚 COURSE MATERIALS

### Primary Text

-   **R for Data Science (1st Edition)** – Free online: <https://r4ds.hadley.nz>

-   **PDF Available**: All 24 chapters included in this plan

### Secondary References

1.  **The Epidemiologist R Handbook** – Free online: <https://www.epirhandbook.com>
    -   Domain-specific examples, epi workflows, case studies
2.  **Advanced R (2nd Edition)** by Hadley Wickham – Free online: <https://adv-r.hadley.nz>
    -   Deep-dive on functions, vectors, environments, metaprogramming

### Required Software

-   **R** (latest version): <https://cloud.r-project.org>
-   **RStudio IDE**: <https://www.rstudio.com/download>

------------------------------------------------------------------------

## 📖 COMPLETE 25-WEEK STUDY SCHEDULE (REORDERED FOR LEARNING)

```{r, include=FALSE}
pacman::p_load(knitr,kableExtra, dplyr)
```

```{r, echo=FALSE}
schedule <- data.frame(
  Session = 1:25,
  Date = c("Feb 4", "Feb 11", "Feb 18", "Feb 25", "Mar 4", "Mar 11", "Mar 18", "Mar 25", 
           "Apr 1", "Apr 8", "Apr 15", "Apr 22", "Apr 29", "May 6", "May 13", "May 20", 
           "May 27", "Jun 3", "Jun 10", "Jun 17", "Jun 24", "Jul 1", "Jul 8", "Jul 15", "Jul 22"),
  Ch = c("2", "4", "6", "7", "8", "9", "3 (Pt 1)", "3 (Pt 2)", "14", "10", "11", "12", 
         "13", "5", "1", "15", "16", "17", "18", "19", "20", "21", "22", "23", "24"),
  Chapter = c("Workflow: Basics", "Workflow: Scripts", "Workflow: Projects", "Tibbles with tibble", 
              "Data Import with readr", "Tidy Data with tidyr", "Data Transformation Pt 1", 
              "Data Transformation Pt 2", "Pipes with magrittr", "Relational Data", 
              "Strings with stringr", "Factors with forcats", "Dates & Times with lubridate",
              "Exploratory Data Analysis", "Data Visualization with ggplot2", "Functions", 
              "Vectors", "Iteration with purrr", "Model Basics with modelr", 
              "Model Building", "Many Models with purrr & broom", "R Markdown", 
              "Graphics for Communication", "R Markdown Formats", "R Markdown Workflow"),
  Key_Topics = c("Coding fundamentals", "Running code, debugging", "RStudio projects", 
                 "Tibbles vs data.frame", "CSV/Excel parsing", "pivot_longer/wider", 
                 "filter, arrange, select", "mutate, summarise, group_by", 
                 "%>% operator, piping", "Keys, joins", "Regex matching", 
                 "Recoding factors", "Time spans", "Patterns, outliers", 
                 "Aesthetics, geoms", "Scoping, arguments", "Atomic vectors", 
                 "map, walk, reduce", "Simple models", "Case studies", 
                 "List columns", "Code chunks, YAML", "Labels, themes", 
                 "Notebooks", "Best practices"),
  Epi_Ref = c("3: R basics", "3: R basics", "3: R basics", "5: Data wrangling", 
              "9: Import/export", "5: Data wrangling", "4: dplyr basics", 
              "4: dplyr basics", "4: dplyr basics", "5: Data wrangling", 
              "8: Character strings", "6: Factors", "7: Dates", 
              "10: Descriptive tables", "11: ggplot2", "14: Functions", 
              "13: Logical vectors", "15: Loops", "20: Statistical modeling", 
              "20: Statistical modeling", "20: Statistical modeling", 
              "18: Output formats", "11: ggplot2", "18: Output formats", 
              "18: Output formats"),
  Adv_R_Ref = c("Ch5: Style", "Ch5: Style", "Ch5: Style", "Ch2: Data structures", 
                "Ch3: Subsetting", "Ch3: Subsetting", "Ch3: Subsetting", 
                "Ch11: Functionals", "Ch11: Functionals", "Ch3: Subsetting", 
                "Ch6: Functions", "Ch2: Data structures", "Ch6: Functions", 
                "Ch5: Style guide", "Ch6: Functions", "Ch6: Functions", 
                "Ch2: Data structures", "Ch11: Functionals", "Ch6: Functions", 
                "Ch16: Performance", "Ch11: Functionals", "Ch5: Style guide", 
                "Ch6: Functions", "Ch5: Style guide", "Ch5: Style guide"),
  Leaders = c("#1", "#2", "#3", "#4", "#5", "#6", "#7", "#8", 
              "#9 & #10", "#11", "#12", "#13", "#14", "#1", "#2", "#3", 
              "#4", "#5 & #6", "#7", "#8", "#9", "#11 & #12", "#10", 
              "#13", "#14")
)

kable(schedule, escape = FALSE, booktabs = TRUE) %>%
  kable_styling(bootstrap_options = c("striped", "hover")) %>%
  row_spec(c(9,18,22), bold = TRUE)
```

## 📚 BOOK STRUCTURE: ALL 24 CHAPTERS

### **PART I: EXPLORE (Chapters 1–6)**

Foundational concepts and workflows before diving into data manipulation.

```{r, echo=FALSE}
part1 <- data.frame(
  Ch = c("1", "2", "3", "4", "5", "6"),
  Title = c("Data Visualization with ggplot2", 
            "Workflow: Basics", 
            "Data Transformation with dplyr", 
            "Workflow: Scripts", 
            "Exploratory Data Analysis", 
            "Workflow: Projects"),
  `Key Sections` = c("First steps, aesthetics, facets, geoms, stats, layered grammar",
                    "Coding basics, naming, functions, console interaction",
                    "filter, arrange, select, mutate, summarise, group_by",
                    "Running code, diagnostics, debugging in RStudio",
                    "Variation, covariation, patterns, models",
                    "RStudio projects, organization, file structure"),
  Week = c("15", "1", "7–8", "2", "14", "3"),
  Competency = c("Create exploratory & publication plots",
                "Understand R fundamentals",
                "Manipulate data rows/columns",
                "Write and execute scripts",
                "Discover insights before modeling",
                "Organize analyses professionally")
)

kable(part1, escape = FALSE, booktabs = TRUE) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover")) %>%
  column_spec(c(1, 4), bold = TRUE)
```
------------------------------------------------------------------------

### **PART II: WRANGLE (Chapters 7–13)**

Data import, cleaning, reshaping, and handling special data types.

```{r, echo=FALSE}
part2 <- data.frame(
  Ch = c("7", "8", "9", "10", "11", "12", "13"),
  Title = c("Tibbles with tibble", 
            "Data Import with readr", 
            "Tidy Data with tidyr", 
            "Relational Data with dplyr", 
            "Strings with stringr", 
            "Factors with forcats", 
            "Dates & Times with lubridate"),
  `Key Sections` = c("Creating, printing, subsetting tibbles",
                    "Parsing, reading CSV/Excel, writing files",
                    "pivot_longer, pivot_wider, separate, missing",
                    "Keys, joins (left/inner/anti), set operations",
                    "String basics, regex, pattern matching, tools",
                    "Creating, ordering, recoding factors",
                    "Creating, components, time spans, zones"),
  Week = c("4", "5", "6", "10", "11", "12", "13"),
  Competency = c("Work with tidy data structures",
                "Import data into R",
                "Reshape messy data",
                "Combine multiple datasets",
                "Handle text data",
                "Manage categorical variables",
                "Handle temporal data")
)

kable(part2, escape = FALSE, booktabs = TRUE) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover")) %>%
  column_spec(c(1, 4), bold = TRUE)
```
------------------------------------------------------------------------

### **PART III: PROGRAM (Chapters 14–17)**

Writing functions and iterating efficiently.

```{r, echo=FALSE}
part3 <- data.frame(
  Ch = c("14", "15", "16", "17"),
  Title = c("Pipes with magrittr", 
            "Functions", 
            "Vectors", 
            "Iteration with purrr"),
  `Key Sections` = c("%>% , piping alternatives, when to use",
                    "Components, scoping, arguments, returns",
                    "Atomic vectors, lists, attributes, augmented",
                    "For loops, map, walk, reduce, functionals"),
  Week = c("9", "16", "17", "18"),
  Competency = c("Chain operations fluently",
                "Write reusable functions",
                "Understand R data types",
                "Loop efficiently")
)

kable(part3, escape = FALSE, booktabs = TRUE) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover")) %>%
  column_spec(c(1, 4), bold = TRUE)
```
------------------------------------------------------------------------

### **PART IV: MODEL (Chapters 18–20)**

Building and evaluating statistical models.

```{r, echo=FALSE}
part4 <- data.frame(
  Ch = c("18", "19", "20"),
  Title = c("Model Basics with modelr", 
            "Model Building", 
            "Many Models with purrr & broom"),
  `Key Sections` = c("Simple models, visualization, formulas, families",
                    "Why models, case studies, evaluation",
                    "List columns, nested data, tidying results"),
  Week = c("19", "20", "21"),
  Competency = c("Understand modeling workflow",
                "Build & evaluate models",
                "Fit models to grouped data")
)

kable(part4, escape = FALSE, booktabs = TRUE) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover")) %>%
  column_spec(c(1, 4), bold = TRUE)
```
------------------------------------------------------------------------

### **PART V: COMMUNICATE (Chapters 21–24)**

Documenting and sharing results reproducibly.

```{r, echo=FALSE}
part5 <- data.frame(
  Ch = c("21", "22", "23", "24"),
  Title = c("R Markdown", 
            "Graphics for Communication", 
            "R Markdown Formats", 
            "R Markdown Workflow"),
  `Key Sections` = c("Basics, formatting, code chunks, YAML",
                    "Labels, annotations, scales, themes, saving",
                    "Documents, notebooks, presentations, dashboards",
                    "Organization, practices, reproducibility"),
  Week = c("22", "23", "24", "25"),
  Competency = c("Write reproducible reports",
                "Polish plots for publication",
                "Create multiple output formats",
                "Apply best practices")
)

kable(part5, escape = FALSE, booktabs = TRUE) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover")) %>%
  column_spec(c(1, 4), bold = TRUE)
```

------------------------------------------------------------------------

## 🔗 SECONDARY REFERENCE MAPPING

Use these resources to clarify concepts and explore domain-specific examples:

```{r, echo=FALSE}
ref_map <- data.frame(
  `R4DS Chapters` = c("2, 4, 6", "7–9", "10–13", "1, 5, 14", "15–17", "18–20", "21–24"),
  `Epidemiologist R Handbook` = c("3: R basics", 
                                 "5: Data wrangling, 9: Import/export", 
                                 "4: dplyr, 6: Factors, 7: Dates, 8: Strings", 
                                 "10: Descriptive tables, 11: ggplot2", 
                                 "14: Functions, 15: Loops", 
                                 "20: Statistical modeling, 21: Models", 
                                 "18: Output formats, 22: Communication"),
  `Advanced R` = c("Ch5: Style guide", 
                   "Ch2, Ch3: Data structures", 
                   "Ch3: Subsetting, Ch6: Functions", 
                   "Ch6: Functions", 
                   "Ch6: Functions, Ch11: Functionals", 
                   "Ch16: Performance", 
                   "Ch5: Style guide"),
  `When to Use` = c("Fundamentals & code organization",
                   "Understanding data manipulation",
                   "Complex transformations",
                   "Visualization & exploration",
                   "Programming concepts",
                   "Modeling & optimization",
                   "Documentation & sharing")
)

kable(ref_map, escape = FALSE, booktabs = TRUE) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover")) %>%
  column_spec(c(1), bold = TRUE)
```
------------------------------------------------------------------------

## 📖 BOOK ORDER vs. OUR LEARNING SEQUENCE

### **Why We're Reordering the Chapters**

The **R for Data Science book is published in this order**: Visualization (Ch 1) → Transformation (Ch 3) → Wrangling (Ch 7–13) → Programming (Ch 14–17) → Modeling (Ch 18–20) → Communication (Ch 21–24).

However, **this order isn't optimal for beginners learning epidemiology**. The book's authors acknowledge this in their Preface:

> "Starting with data ingest and tidying is suboptimal because 80% of the time it's routine and boring, and the other 20% of the time it's weird and frustrating. That's a bad place to start learning a new subject! Instead, we'll start with visualization and transformation of data that's already been imported and tidied."

**The problem:** Chapter 1 (ggplot2) assumes clean data exists. But beginners need to **import, clean, and reshape data FIRST** before they can visualize or analyze it.

### **Our Reordered Learning Sequence**

We follow the **natural real-world epidemiological workflow** instead of the book's ordering:

```{r, echo=FALSE}
sequence <- data.frame(
  Phase = c("1. Foundation", "2. Wrangling", "3. Exploration", 
            "4. Programming", "5. Modeling", "6. Communication"),
  Weeks = c("1–3", "4–13", "14–15", "16–18", "19–21", "22–25"),
  `Our Order` = c("First", "Second", "Third", "Fourth", "Fifth", "Sixth"),
  `Book Chapters` = c("2, 4, 6 (Workflow)", 
                     "7–13 (Import, Tidy, Transform, Join, Handle types)", 
                     "5, 1 (EDA, Visualization)", 
                     "15, 16, 17 (Functions, Vectors, Iteration)", 
                     "18, 19, 20 (Model Basics, Building, Many Models)", 
                     "21–24 (R Markdown, Graphics, Formats, Workflow)"),
  Purpose = c("Learn RStudio, organize work, set up projects BEFORE coding",
             "Get messy data into analysis-ready format",
             "Explore clean data visually & statistically",
             "Write reusable code & automate workflows",
             "Fit & evaluate statistical models",
             "Document findings & create reproducible reports")
)

kable(sequence, escape = FALSE, booktabs = TRUE) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover")) %>%
  column_spec(c(2), bold = TRUE)
```

### **Why This Order Works Better**

✅ **Motivation**: Learn foundational tools first, then apply them to real problems\
✅ **Real-world**: Follow the actual epidemiological analysis workflow\
✅ **Practical**: Work with messy data early; clean it before analyzing\
✅ **Confidence**: See tangible results at each step (import → clean → analyze → visualize)\
✅ **Less frustration**: Don't start with "boring" wrangling; build to it naturally

### **Book Order (As Published)**

For reference, here's the original chapter sequence in R for Data Science (1st Edition):

```{r, echo=FALSE}
book_order <- data.frame(
  Ch = c("1", "2", "3", "4", "5", "6", "7–13", "14–17", "18–20", "21–24"),
  Part = c("I: Explore", "I: Explore", "I: Explore", "I: Explore", 
           "I: Explore", "I: Explore", "II: Wrangle", "III: Program", 
           "IV: Model", "V: Communicate"),
  Title = c("Data Visualization with ggplot2",
            "Workflow: Basics",
            "Data Transformation with dplyr",
            "Workflow: Scripts",
            "Exploratory Data Analysis",
            "Workflow: Projects",
            "Tibbles, Import, Tidy, Joins, Strings, Factors, Dates",
            "Pipes, Functions, Vectors, Iteration",
            "Model Basics, Building, Many Models",
            "R Markdown, Graphics, Formats, Workflow")
)

kable(book_order, escape = FALSE, booktabs = TRUE) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover")) %>%
  column_spec(c(1), bold = TRUE)
```

------------------------------------------------------------------------

## 🎓 WEEKLY MEETING FORMAT (60 Minutes)

### **Structure**

```{r, echo= FALSE}
meeting_format <- data.frame(
  Time = c("0–15 min", "15–45 min", "45–55 min", "55–60 min"),
  Activity = c("Summary & Discussion", 
               "Live Demo, Exercises & Problem-Solving", 
               "Help & Troubleshooting", 
               "Wrap-up & Next Week Preview"),
  Details = c("Leader(s) summarize key concepts & learning objectives; group discusses what was clear/confusing and what's relevant to epi work",
             "Leader(s) demonstrate 2–3 code examples live; group follows along, attempts exercises, asks questions in real-time",
             "Leader(s) & peers help resolve coding errors, confusions, and remaining questions",
             "Summarize key takeaways; preview next week's chapter; assign prep work")
)

kable(meeting_format, escape = FALSE, booktabs = TRUE) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover")) %>%
  column_spec(c(1), bold = TRUE)
```

### **Leader Responsibilities**

Each leader should: 
- **Read chapter thoroughly** 1 week in advance (3–4 hours) 
- **Prepare 2–3 worked code examples** to demo live (copy-paste ready) 
- **Identify 3–5 key concepts** that will be discussed 
- **Note common confusions** or tricky sections from your own reading 
- **Bring 2–3 discussion questions** to engage the group 
- **Help troubleshoot** participant questions & coding errors 
- **Share practical epi applications** of the chapter content

**For Co-Leaders (3 toughest sessions):** 
- **Divide responsibilities**: One leads first half, other leads second half  
- **Coordinate prep work**: Meet 1 week before to align on content & demos 
- **Share code examples**: Create demo scripts together for consistency

### **Leadership Rotation (Random Selection by Number 1–14)**

**To assign leaders randomly:** 
1. Each person selects a number between **1 and 14** at the start of the course 
2. Leadership follows the rotation below across all 25 sessions

```{r, echo=FALSE}
rotation <- data.frame(
  Number = c("#1", "#2", "#3", "#4", "#5", "#6", "#7", "#8", "#9", "#10", "#11", "#12", "#13", "#14"),
  `Session 1` = c("Feb 4", "Feb 11", "Feb 18", "Feb 25", "Mar 4", "Mar 11", "Mar 18", "Mar 25", 
                  "Apr 1(co)", "Apr 1(co)", "Apr 8", "Apr 15", "Apr 22", "Apr 29"),
  `Session 2` = c("May 6", "May 13", "May 20", "May 27", "Jun 3(co)", "Jun 3(co)", "Jun 10", 
                  "Jun 17", "Jun 24", "Jul 8", "Jul 1(co)", "Jul 1(co)", "Jul 15", "Jul 22"),
  `Total` = rep("2 sessions", 14)
)

kable(rotation, escape = FALSE, booktabs = TRUE) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover")) %>%
  column_spec(c(1), bold = TRUE)
```

**Toughest sessions (2 leads each):**
Apr 1 (Session 9): #9 & #10 co-lead (Pipes - Ch 14)
Jun 3 (Session 18): #5 & #6 co-lead (Iteration - Ch 17)
Jul 1 (Session 22): #11 & #12 co-lead (R Markdown - Ch 21)
Perfect balance: 14 leads × 2 sessions each = exactly 25 sessions covered!

------------------------------------------------------------------------

## 📝 WEEKLY PREPARATION (~**4 hours**)

Participants should dedicate approximately **4 hours per week**:

### **Before the Meeting** (3.5 hours)

1.  **Read & understand chapter** (**2 hours**)
    - Skim introduction for context (15 min)
    - Read thoroughly with code examples (1 hr)
    - Run **all** examples in RStudio (45 min)  
    - Take notes on key concepts (15 min)

2.  **Practice exercises** (**1 hour**)
    - Work through **5–7 exercises** at chapter end
    - Run code locally; save working examples
    - Debug errors independently first
    - Note 2–3 challenges for discussion

3.  **Prepare questions & review** (**30 minutes**)
    - List **2–3 specific questions** 
    - Review epi applications from syllabus
    - Preview next week's chapter overview

### **After the Meeting** (**30 minutes**)
- Review meeting notes & code demos
- Re-run 1–2 examples from leader
- Update personal code library

------------------------------------------------------------------------

## 🔬 EPIDEMIOLOGICAL APPLICATIONS

See how each chapter applies to real STH research:

```{r, echo=FALSE}
epi_apps <- data.frame(
  Chapter = c("8", "9", "3, 10", "11", "12, 13", "5, 1", "15–17", "18–20", "21–24"),
  `Epi Application` = c("Import surveillance data",
                       "Reshape prevalence data",
                       "Filter & join datasets",
                       "Parse case definitions",
                       "Manage categories & timelines",
                       "Visualize distribution",
                       "Automate calculations",
                       "Model transmission",
                       "Generate reports"),
  Example = c("Load district-level STH prevalence data from Excel",
             "Pivot age-stratified prevalence wide to long format",
             "Subset by age group; join with population denominators",
             "Extract infection intensity category from text",
             "Organize treatment arms; track follow-up dates",
             "Create maps & plots of STH by region & year",
             "Write function to calculate prevalence for any region",
             "Fit age-prevalence curves; predict MDA impact",
             "Create quarterly surveillance reports in HTML/PDF")
)

kable(epi_apps, escape = FALSE, booktabs = TRUE) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover")) %>%
  column_spec(c(1), bold = TRUE)
```

------------------------------------------------------------------------

## 🤝 GROUP GUIDELINES

### **Expectations**

-   **Attend weekly meetings**: In-person or virtual, same time each week
-   **Complete prep work**: Read chapter & attempt exercises before meeting
-   **Participate actively**: Share insights, ask questions, help peers
-   **Lead 2–3 chapters**: Fulfill your leadership duties when assigned (solo or co-lead)
-   **Respect others**: Patient with different learning paces & backgrounds

### **Communication**

-   **WhatsApp channel**: Share resources, ask quick questions between meetings
-   **Google Drive folder**: Store notes, code, datasets
-   **Meeting notes**: Record key takeaways & useful code snippets each week

------------------------------------------------------------------------

## 📞 SUPPORT & RESOURCES

### **If You Get Stuck**

1.  **Google the error message** – Often finds solutions quickly
2.  **Search Stack Overflow** – Add `[R]` tag for R-specific results
3.  **Consult secondary references**:
    -   Epidemiologist R Handbook: <https://www.epirhandbook.com>
    -   Advanced R: <https://adv-r.hadley.nz>
4.  **Ask the group**: Bring confusing sections to weekly meetings
5.  **WhatsApp channel**: Post quick questions between meetings

### **Helpful Links**

-   **R Project**: <https://www.r-project.org>
-   **RStudio Download**: <https://www.rstudio.com>
-   **Tidyverse**: <https://www.tidyverse.org>
-   **R4DS Online**: <https://r4ds.hadley.nz>
-   **RStudio Community**: <https://community.rstudio.com>
-   **Stack Overflow `[R]` tag**: <https://stackoverflow.com/questions/tagged/r>

------------------------------------------------------------------------

## 📅 TIMELINE SUMMARY

```{r, echo=FALSE}
timeline <- data.frame(
  Phase = c("Foundations", "Wrangling", "Programming", "Modeling", "Communication"),
  Weeks = c("1–3", "4–13", "14–18", "19–21", "22–25"),
  Chapters = c("2, 4, 6", "7–13", "5, 1, 14–17", "18–20", "21–24"),
  Focus = c("RStudio, scripts, organization",
           "Import, clean, reshape, join data",
           "EDA, visualization, functions, iteration",
           "Fit & evaluate models",
           "Reports, dashboards, reproducibility")
)

kable(timeline, escape = FALSE, booktabs = TRUE) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover")) %>%
  column_spec(c(3), bold = TRUE)
```
------------------------------------------------------------------------

## ✅ CHECKLIST FOR GETTING STARTED

Before Week 1 (by February 04, 2026):

-   [ ] Install R (latest version)
-   [ ] Install RStudio IDE
-   [ ] Test RStudio: Open it, run `1 + 1` in console
-   [ ] Read Chapter 2 (Workflow: Basics) for first meeting (Leader #1)

------------------------------------------------------------------------

## 📖 FINAL NOTES

This syllabus provides a **structured path to mastery** of R for epidemiological data science. The key to success is:

1.  **Consistency**: Show up weekly & do the prep work
2.  **Curiosity**: Ask questions; try variations of examples
3.  **Collaboration**: Learn from peers; help each other
4.  **Application**: Apply concepts to your own epi data

By **Week 25**, you'll have the skills to conduct independent, reproducible epidemiological analyses from raw data to publication-ready reports.

------------------------------------------------------------------------

## 📝 Document Version Control

```{r, echo=FALSE}
versions <- data.frame(
  Version = c("1.0", "2.0"),
  Date = c("Jan 21, 2026", "Jan 29, 2026"),
  Changes = c("Initial complete syllabus with all 24 chapters, 25 weeks, 10-person rotation, 60-min meetings, co-leaders in final 5 weeks, chapters reordered for learning progression",
             "Updated for 14 leads (each gets exactly 2 sessions), start date Feb 4, 3 toughest sessions with co-leaders (Pipes, Iteration, R Markdown), date-based session schedule")
)

kable(versions, escape = FALSE, booktabs = TRUE) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover"))
```

------------------------------------------------------------------------

*Last updated: **January 29, 2026** | Next review: **March 2026***

**Ready to get started? Let's master R for epidemiology! 🚀**


<h2>Get your lead number</h2>
<p>Type your full name to generate a lead number (1–14). Please do not generate a number more than once for yourself. If you use the same name again on this device/browser, you’ll get the same number (incase you have forgotten you lead number).</p>

<div id="lead-generator" style="max-width: 820px;">
  <label for="leadName"><strong>Your name</strong></label><br>
  <input id="leadName" type="text" placeholder="e.g., Gideon J I" style="width: 100%; padding: 8px; margin: 6px 0;">
  <button id="leadBtn" style="padding: 8px 12px;">Generate</button>
  <button id="leadResetBtn" style="padding: 8px 12px; margin-left: 8px;">Admin reset</button>
  <p id="leadMsg" style="margin-top: 10px;"></p>

  <h3>Assigned leads (this browser)</h3>
  <table style="width:100%; border-collapse: collapse;" border="1" cellpadding="6">
    <thead>
      <tr>
        <th>#</th>
        <th>Name</th>
        <th>Lead number</th>
        <th>Session 1</th>
        <th>Session 2</th>
        <th>Total</th>
      </tr>
    </thead>
    <tbody id="leadTableBody"></tbody>
  </table>

  <p id="leadFullMsg" style="margin-top:10px;"></p>
</div>